Vaccines: Many Beliefs Study 3: Analysis

### 2018-11-29 study1/custom-structure-learning-vaccData.Rmd
before yale visit

# source("cog-model-graph-tools.R")
source("../scripts/custom-structure-learning/cog-model-main.R", chdir = TRUE)
source("../scripts/gmod_tools.R")
source("../scripts/bnTheoryBlacklist.R")

@@

```{r}
# takes about 20-30 min
start_time <- Sys.time()

# test2 <- test
predictions <- test %>% mutate_all(function(x){NA}) %>% mutate_all(as.numeric)
# net.fit <- bn.fit(map_dag, train)

for (targVar in names(test)) {
  tempDF <- test
  tempDF[[targVar]] <- NA
  tempDF[[targVar]] <- as.numeric(tempDF[[targVar]])

  # imputed <- impute(net.fit, tempDF, method="bayes-lw")
  imputed <- make_predictions(model_string, test, nodes_to_predict = targVar)
  predictions[[targVar]] <- imputed[[targVar]]
}

predictions <- predictions %>% gather(varName, predValue)
test_analyze <- test %>% gather(varName, value) %>% bind_cols(predictions) %>% select(varName, value, predValue)

print(Sys.time() - start_time)
```

@@

@@

corrDF <- test_analyze %>%
  group_by(varName) %>%
  summarize(cor = cor(value,predValue)) %>%
  mutate(cor = format(round(cor,2),nsmall=2))

plt.valid <- test_analyze %>%
  ggplot(aes(x=predValue, y=value)) + 
  geom_abline(slope=1, intercept=0, linetype="dashed", alpha=.33) +
  geom_point(shape=1, size=.5, color="darkturquoise") + 
  geom_text(data=corrDF, aes(label=paste("r =", cor)),
            x=-Inf, y=Inf, hjust=-0.12, vjust=1.4,
            size = 10*.352777778,
            color="grey25") +
  # geom_smooth(method="lm") +
  coord_fixed() + 
  facet_wrap(~varName, nrow=2) + 
  scale_x_continuous("Predicted values", limits = c(0, 1)) +
  scale_y_continuous("Observed values", limits = c(0, 1)) +
  theme_bw(base_size = 10) +
  theme(strip.background = element_rect(fill="grey90"), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

plt.valid
```

>>>>

# source("cog-model-graph-tools.R")
source("../scripts/custom-structure-learning/cog-model-main.R", chdir = TRUE)
source("../scripts/gmod_tools.R")
source("../scripts/bnTheoryBlacklist.R")
source("../scripts/custom-structure-learning/cog-model-jags-tools.R")

@@

```{r}
# #  takes about 20-30 min to run

# start_time <- Sys.time()
# 
# # test2 <- test
# predictions <- test %>% mutate_all(function(x){NA}) %>% mutate_all(as.numeric)
# # net.fit <- bn.fit(map_dag, train)
# 
# for (targVar in names(test)) {
#   tempDF <- test
#   tempDF[[targVar]] <- NA
#   tempDF[[targVar]] <- as.numeric(tempDF[[targVar]])
#   
#   # imputed <- impute(net.fit, tempDF, method="bayes-lw")
#   imputed <- make_predictions(model_string, test, nodes_to_predict = targVar)
#   predictions[[targVar]] <- imputed[[targVar]]
# }
# 
# predictions <- predictions %>% gather(varName, predValue)
# test_analyze <- test %>% gather(varName, value) %>% bind_cols(predictions) %>% select(varName, value, predValue)
# 
# print(Sys.time() - start_time)

# saveRDS(object = test_analyze, file = "../local/s1_model_predictions.rds")

test_analyze <- readRDS("../local/s1_model_predictions.rds")
```

# job talk plots

November 16, 2018

## predicted vs observed across scales

@@

recode_nodes <- function(x){
  recode(x,
         "diseaseRare" = "Disease Rarity",
         "diseaseSevere" = "Disease Severity",
         "hb" = "Holistic Balance",
         "infantImmLimCap" = "IIS: Limited\nCapacity",
         "infantImmWeak" = "IIS: Weakness",
         "medSkept" = "Medical\nSkepticism",
         "nat" = "Naturalism",
         "overpar" = "Parental\nProtectiveness",
         "parentExpert" = "Parental\nExpertise",
         "vaccDanger" = "Vaccine\nDanger",
         "vaccEff" = "Vaccine\nEffectiveness",
         "vaccIntent" = "Vaccination\nIntentions",
         "vaccStrain" = "IIS: Vaccines\nStrain",
         "vaccTox" = "Vaccines\nToxicity")}

@@

corrDF <- test_analyze %>%
  # mutate(varName = recode_nodes(varName)) %>%
  group_by(varName) %>%
  summarize(cor = cor(value,predValue)) %>%
  mutate(cor = format(round(cor,2),nsmall=2))

plt.valid <- test_analyze %>%
  mutate(varName = recode_nodes(varName)) %>%
  ggplot(aes(x=predValue, y=value)) + 
  geom_abline(slope=1, intercept=0, linetype="dashed", alpha=.33) +
  geom_point(shape=1, size=.5, color="darkturquoise") + 
  geom_text(data=corrDF, aes(label=paste("r =", cor)),
            x=-Inf, y=Inf, hjust=-0.12, vjust=1.4,
            size = 10*.352777778,
            color="grey25") +
  # geom_smooth(method="lm") +
  coord_fixed() + 
  facet_wrap(~varName, nrow=2) + 
  scale_x_continuous("Predicted values", limits = c(0, 1)) +
  scale_y_continuous("Observed values", limits = c(0, 1)) +
  theme_bw(base_size = 10) +
  theme(strip.background = element_rect(fill="grey90"), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggsave(plot = plt.valid, filename = "../local/s1_predobs_plot.png", width=8.5, height=4)
```

## network structure (beautified)

```{r}
# do network structure plotting here ...
library(DiagrammeR)

custom_plot.HydeNetwork <- function(x, 
                             customNodes = NULL,
                             customEdges = NULL,
                             ..., 
                             removeDeterm = FALSE,
                             useHydeDefaults = TRUE)
{
  if (removeDeterm) x <- plot_nondeterm_only(x)

  node_df <- 
    DiagrammeR::create_node_df(n = length(x[["nodes"]]),
                               label = x[["nodes"]])
  # 
  # node_df <- data.frame(nodes = x[["nodes"]],
  #                       stringsAsFactors = FALSE)
  if (useHydeDefaults) node_df <- HydeNet:::mergeDefaultPlotOpts(x, node_df)

  if (!is.null(customNodes)) node_df <- HydeNet::mergeCustomNodes(node_df, customNodes)

  edge_table <- do.call("rbind", 
                        mapply(FUN = HydeNet:::mapEdges, 
                               x[["nodes"]], 
                               x[["parents"]],
                               MoreArgs = list(node_df = node_df)))

  edge_df <- DiagrammeR::create_edge_df(from = edge_table[, 2], 
                                        to = edge_table[, 1])

  if (!is.null(customEdges)) HydeNet::mergeCustomEdges(edge_df, customEdges)

  DiagrammeR::create_graph(nodes_df = node_df,
                           edges_df = edge_df,
                           attr_theme = NULL)

}

## ----------------------------------------

# code to demo. For some reason set_edge_attrs() isn't working right, so have to do it manually

map_to_viridis <- function(vec, bins=64, option="D"){

  binned <- cut(vec, bins)
  levels(binned) <- viridis::viridis(bins, option=option)

  return(as.character(binned))
}

map_to_brewer <- function(vec, bins=11, option="RdYlBu"){
  binned <- cut(vec, bins)
  levels(binned) <- RColorBrewer::brewer.pal(bins, name=option)

  return(as.character(binned))
}

```{r}
z <- custom_plot.HydeNetwork(h)

c <- find_dag_coefs(map_dag, train)

coefs <- tidy_coefs(c)

coefs %>%
  filter(parameter!="phi",parameter!="w_leak") %>%
  rename(from = parameter) %>%
  mutate(from = gsub("w_","",from)) %>%
  select(to, from, value) %>%
  merge(z$nodes_df %>% rename(from = label) %>% select(id, from), by="from") %>%
  mutate(from = id) %>%
  select(-id) %>%
  merge(z$nodes_df %>% rename(to = label) %>% select(id, to), by="to") %>%
  mutate(to = id) %>%
  arrange(to,from) -> edge_df_mod
  # select(value, id.x, id.y) %>%
  # mutate(from = id.x, to = id.y)

# z$nodes_df
z$nodes_df <- z$nodes_df %>% mutate(label = recode_nodes(label))
z$nodes_df$fontname <- "Helvetica"
z$nodes_df$shape <- "ellipse"
z$nodes_df$fillcolor <- ifelse(z$nodes_df$id==13, "grey","white")
z$nodes_df$fixedsize <- FALSE
z$edges_df$penwidth <- abs(edge_df_mod$value)*2
z$edges_df$penwidth <- 1.5
# z$edges_df$label <- round(edge_df_mod$value,2)
# z$edges_df$color <- ifelse(edge_df_mod$value > 0, "blue","red")
z$edges_df$color <- map_to_brewer(edge_df_mod$value)
render_graph(z)

# save relevant outputs!

```{r}
# save model as bn
saveRDS(map_dag, "../local/map_dag_bn_theorybased.rds")
# and as hydnet
saveRDS(h, "../local/map_dag_hydenet_theorybased.rds")
# save model string
saveRDS(model_string, "../local/map_dag_modelstring_theorybased.rds")

### 2018-11-29 study2/vacc_manyBelief_study2_analysis.Rmd
refactored, working
before yale visit

```{r tidy, include = F}
# load data
d_all <- read.csv("../study3/data/study3_data.csv")[-1]

@@

```{r plot diffs, fig.width = 3, fig.asp = 3}
d_diffs <- d_scored %>%
  spread(phase, mean) %>%
  mutate(post_pre_diff = post - pre) %>%
  group_by(condition, scale) %>%
  do(data.frame(rbind(smean.cl.boot(.$post_pre_diff)))) %>%
  ungroup() %>%
  mutate(scale = factor(scale,
                        levels = c("vaccIntent", "vaccDanger", "vaccEff", 
                                   "vaccStrain", "vaccTox", 
                                   "diseaseSevere", "diseaseRare", 
                                   "infantImmLimCap", "infantImmWeak", 
                                   "medSkept", "hb", "nat", 
                                   "overpar", "parentExpert")),
         condition = factor(condition,
                            levels = c("diseaseRisk", "autismCorr", "noInterv")))

g_diffs <- ggplot(d_diffs,
                  aes(x = condition, y = Mean, 
                      color = condition, group = condition)) +
  facet_wrap(~ scale, ncol = 3) +
  geom_hline(yintercept = 0, lty = 2, size = 0.5) +
  geom_linerange(aes(ymin = Lower, ymax = Upper)) +
  geom_point(size = 3) +
  scale_y_continuous("mean diff") + #, 
                     # limits = c(-3, 3), 
                     # breaks = seq(-3, 3, 1)) +
  scale_color_brewer(palette = "Set1", direction = -1) +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "mean difference scores by scale and condition",
       subtitle = "error bars are 95% bootstrapped CIs")

>>>>

```{r tidy, include = F}
# load data
d_all <- read.csv("../data/study2_data.csv")[-1]

@@

```{r plot diffs}
d_diffs <- d_scored %>%
  spread(phase, mean) %>%
  mutate(post_pre_diff = post - pre) %>%
  group_by(condition, scale) %>%
  do(data.frame(rbind(smean.cl.boot(.$post_pre_diff)))) %>%
  ungroup() %>%
  mutate(scale = factor(scale,
                        levels = c("vaccIntent", "vaccDanger", "vaccEff", 
                                   "vaccStrain", "vaccTox", 
                                   "diseaseSevere", "diseaseRare", 
                                   "infantImmLimCap", "infantImmWeak", 
                                   "medSkept", "hb", "nat", 
                                   "overpar", "parentExpert")),
         condition = factor(condition,
                            levels = c("diseaseRisk", "autismCorr", "noInterv")))

g_diffs <- ggplot(d_diffs,
                  aes(x = scale, y = Mean, 
                      color = condition, group = condition)) +
  # facet_wrap(~ scale, ncol = 3) +
  geom_hline(yintercept = 0, lty = 2, size = 0.5) +
  geom_linerange(aes(ymin = Lower, ymax = Upper)) +
  geom_point(size = 3) +
  scale_y_continuous("mean diff") + #, 
                     # limits = c(-3, 3), 
                     # breaks = seq(-3, 3, 1)) +
  scale_color_brewer(palette = "Set1", direction = -1) +
  theme_bw() +
  theme(
    # legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "mean difference scores by scale and condition",
       subtitle = "error bars are 95% bootstrapped CIs")
